---
title: "Figures"
author: "Janika Kerner"
output: word_document
---

load libraries

```{r libraries, include=FALSE}
library(data.table) # create empty data frame (data.table)
library(ggplot2)
library(grid) # grid png-files to be able to arrange them
library(gridExtra) # arrange png-files
library(magrittr) # pipe operators %>%, %<>%, %T>%
library(mgcv) # GAMs
library(patchwork) # assemble multiple graphs together
library(piecewiseSEM) # calculate SEMs
library(plyr) # revalue factor levels
library(png)
library(qgraph) # plot SEMs nicely
library(tidymv) # GAM predictions
library(tidyverse)
library(VennDiagram) # venn diagrams
```

```{r set-up, include = FALSE}
# Create objects for easy access to the folders related to the project
data_proc <- file.path("results", "data", "processed")
figures_dir <- file.path("text", "figures")
suppl_dir <- file.path("text", "supplementary")
```

read data

```{r loaddata, include=FALSE}
## plot data ##
pd<-read.csv(here::here(data_proc, "plotdata.csv"), sep=";")
pd$plot<-as.factor(pd$plot)

# transformation of not normally distributed annual temperature
pd$tempy<-log(max(pd$tempy+1)-pd$tempy); pd$tempy<-max(pd$tempy)-pd$tempy
```

# Main text
## 1. Venn diagrams

```{r venndiagram}
slmal<-read.csv(here::here(data_proc, "speclistmal.csv"), row.names = 1)
sldip<-read.csv(here::here(data_proc, "speclistdip.csv"), row.names = 1)

# pollinators
venn.diagram(
  x = list(slmal$spec[slmal$Order!="Orthoptera"], sldip$spec[sldip$Order!="Orthoptera"]),
  category.names = c("malaise traps" , "transect walks"),
  filename = here::here(figures_dir, "venns-single", "venndiagpoll.png"),
  output = TRUE,
          imagetype="png" ,
          height = 2000, 
          width = 2000, 
          resolution = 600,
          col= NA,
          fill = c(alpha("#21908dff",0.7), alpha("#440154ff",0.7)),
          cex = 1.1,
          fontfamily = "sans",
          cat.cex = 1.1,
          cat.pos = c(340, 20),
          cat.dist = c(0.055, 0.055),
          cat.fontfamily = "sans",
          cat.col = c("#21908dff", "#440154ff"),
          disable.logging =T, # avoid creation of log.txt in figure folder
          print.mode = c("raw", "percent"), # plot species number & percentage
          sigdigs=2 # percent without decimal numbers
        )

# grasshoppers
venny<-function(order, cat.name1, cat.name2, pos1, pos2, rotation, extpos, extdist, extlength, cdist, cexy) {
  venn.diagram(
    x = list(slmal$spec[slmal$Order==order], sldip$spec[sldip$Order==order]),
    category.names = c(cat.name1, cat.name2),
    filename = here::here(figures_dir, "venns-single", paste0("venndiag", order, ".png")),
    output = TRUE,
          imagetype="png" ,
          height = 2000, 
          width = 2000, 
          resolution = 600,
          col=NA,
          fill = c(alpha("#21908dff",0.7), alpha("#440154ff",0.7)),
          cex = cexy,
          fontfamily = "sans",
          cat.cex = cexy,
          cat.pos = c(pos1, pos2),
          cat.dist = c(cdist, cdist),
          cat.fontfamily = "sans",
          cat.col = c("#21908dff", "#440154ff"),
          disable.logging =T, # avoid creation of log.txt in figure folder
          print.mode = c("raw", "percent"), # plot species number & percentage
          sigdigs=c(2,2,2), # percent without decimal numbers
          rotation.degree=rotation,
          ext.pos = extpos, # position of labels outside the circles
          ext.dist = extdist, # shorten line to label outside the circles
          ext.length = extlength
        )
}

venny("Orthoptera", "malaise traps", "transect walks", 340, 20, 0, 60, -0.2, 0, 0.04, 1.3)
```

## 2. Elevational & sampling temperature patterns pollinators & grasshoppers

```{r methpatternsfunc}
mp<-pivot_longer(data=pd, cols=c(clpolldiv, clpolldip, divor, ordip),names_to = c("group"),values_to="spec.no") 
mp$method <-case_when(mp$group %in% c("clpolldiv", "divor") ~ "malaise traps",
            mp$group %in% c("clpolldip", "ordip") ~ "transect walks") # define method
mp$method<-factor(mp$method, ordered = T) # factor needs to be ordered for ANOVA like GAM summary

mp$guild <-case_when(mp$group %in% c("clpolldiv", "clpolldip") ~ "pollinators",
            mp$group %in% c("divor", "ordip") ~ "grasshoppers") # define species groups
mp$guild<-factor(mp$guild, ordered = T)

# adjusting sample temperature to use only sampling periods of each method
mp$tempright <-case_when(mp$group =="clpolldip" ~ mp$samplpoll,
                         mp$group == "ordip" ~ mp$samplor,
                         mp$group %in% c("clpolldiv", "divor") ~ mp$tempsampl)


modpollelev<-gam(spec.no~s(elev, k=4)+s(elev,by=method,k=4)+method, data=mp[mp$guild=="pollinators",], method="REML")
modpollelev<-get_gam_predictions(modpollelev, elev, series_length = (max(mp$elev)-min(mp$elev))+1)
modpollelev$guild<-"pollinators"

modorelev<-gam(spec.no~s(elev, k=4)+s(elev,by=method,k=4)+method, data=mp[mp$guild=="grasshoppers",], method="REML")
modorelev<-get_gam_predictions(modorelev, elev, series_length = (max(mp$elev)-min(mp$elev))+1)
modorelev$guild<-"grasshoppers"

modpolltemp<-gam(spec.no~s(tempright, k=4)+s(tempright,by=method,k=4)+method+s(elev, k=4), data=mp[mp$guild=="pollinators",], method="REML")
modpolltemp<-get_gam_predictions(modpolltemp, tempright, series_length = 100*
                                   (max(mp$tempright[mp$guild=="pollinators"])-min(mp$tempright[mp$guild=="pollinators"]))+1)

modpolltemp<-rbind(modpolltemp[modpolltemp$method=="malaise traps" &
                           modpolltemp$tempright>=min(mp$tempsampl) &
                           modpolltemp$tempright<=max(mp$tempsampl),],
               modpolltemp[modpolltemp$method=="transect walks" &
                           modpolltemp$tempright>=min(mp$samplpoll) &
                           modpolltemp$tempright<=max(mp$samplpoll),])

modortemp<-gam(spec.no~s(tempright, k=4)+s(tempright,by=method,k=4)+s(elev, k=4)+method, data=mp[mp$guild=="grasshoppers",], method="REML")
modortemp<-get_gam_predictions(modortemp, tempright, series_length = 100*
                                   (max(mp$tempright[mp$guild=="grasshoppers"])-min(mp$tempright[mp$guild=="grasshoppers"]))+1)
modortemp$guild<-"grasshoppers"
modortemp<-rbind(modortemp[modortemp$method=="malaise traps" &
                           modortemp$tempright>=min(mp$tempsampl) &
                           modortemp$tempright<=max(mp$tempsampl),],
                 modortemp[modortemp$method=="transect walks" &
                           modortemp$tempright>=min(mp$samplor) &
                           modortemp$tempright<=max(mp$samplor),])


methplot<-function (var, xaxisname, dataframe, guild, buc) {
  ggplot(aes({{var}}, spec.no, group=method, col=method, fill=method), data=dataframe) +
  geom_ribbon(aes(ymin = CI_lower, ymax = CI_upper), alpha=0.3, colour = NA) + # colour=NA to remove outer contours
  geom_point(alpha=0.6, pch=16, data=mp[mp$guild==guild,]) +
  geom_line(lwd=1)+
  theme_classic() +
  theme(axis.text = element_text(size=10),
        axis.title = element_text(size=10),
        strip.text = element_blank()) + # suppress plot labelling by facetwrap
  labs(x=xaxisname, y="species richness") +
  scale_color_manual(values = c("malaise traps"="#21908dff", "transect walks"="#440154ff"),name="sampling\nmethod", labels=c("malaise traps"="malaise\ntraps", "transect walks"="transect\nwalks")) +
    scale_fill_manual(values = c("malaise traps"="#21908dff", "transect walks"="#440154ff"), name="sampling\nmethod", labels=c("malaise traps"="malaise\ntraps", "transect walks"="transect\nwalks")) +
  guides(alpha="none") + # no legend for alpha
    annotate("text", x=-Inf, y=Inf, label=buc, hjust=-1, vjust=1, size=5)
}

# plot only points without function lines for grasshopper sampling temperature (in model not significant, only significant effect of elevation)
methpoints<-function (var, xaxisname, dataframe, guild, buc) {
  ggplot(aes({{var}}, spec.no, group=method, col=method, fill=method), data=dataframe) +
  geom_point(alpha=0.6, pch=16, data=mp[mp$guild==guild,]) +
  theme_classic() +
  theme(axis.text = element_text(size=10),
        axis.title = element_text(size=10),
        strip.text = element_blank()) + # suppress plot labelling by facetwrap
  labs(x=xaxisname, y="species richness") +
  scale_color_manual(values = c("malaise traps"="#21908dff", "transect walks"="#440154ff"),name="sampling\nmethod", labels=c("malaise traps"="malaise\ntraps", "transect walks"="transect\nwalks")) +
    scale_fill_manual(values = c("malaise traps"="#21908dff", "transect walks"="#440154ff"), name="sampling\nmethod", labels=c("malaise traps"="malaise\ntraps", "transect walks"="transect\nwalks")) +
  guides(alpha="none") + # no legend for alpha
    annotate("text", x=-Inf, y=Inf, label=buc, hjust=-1, vjust=1, size=5)
}


png(here::here(figures_dir, "methpatterns.png"), height=15, width=18, res=600, units = "cm")

((methplot(elev, "elevation [m a.s.l.]", modpollelev, "pollinators", "a)") + theme(legend.position="none") +
     annotation_custom(rasterGrob(readPNG(here::here(figures_dir, "logos", "clpolldiv.png")), width = .2, x=.9, y=.92))) +
    
    (methplot(elev, "elevation [m a.s.l.]", modorelev, "grasshoppers", "b)") + theme(legend.position="none") +
     annotation_custom(rasterGrob(readPNG(here::here(figures_dir, "logos", "divor.png")), width = .2, x=.9, y=.92)))) /
  
  ((methplot(tempright, "sampling temperature [°C]", modpolltemp, "pollinators", "c)") +
      theme(legend.position=c(.95, .2), legend.title = element_blank(), legend.text = element_blank(), legend.background = element_blank()) +
     annotation_custom(rasterGrob(readPNG(here::here(figures_dir, "logos", "malaisetrap.png")), width = .14, x=.8, y=.25)) +
     annotation_custom(rasterGrob(readPNG(here::here(figures_dir, "logos", "kescher.png")), width = .14, x=.8, y=.1)) +
     scale_x_continuous(breaks = c(15,20,25))) +
     
     (methpoints(tempright, "sampling temperature [°C]", mp, "grasshoppers", "d)") +
     theme(legend.position="none") + scale_x_continuous(breaks = c(15,20,25))))

dev.off()
```

## 3. & 4. Path models

### functions

```{r pathfunctions}
#### Estimates ####

# manual matrix with estimates from path model for plotting

estimates<-function(v, sampltemp, div, plrich) {
  
  t<-matrix(c(rep(0,6*6)),6,6,byrow=T)

  colnames(t)<-c(sampltemp,"tempy","pp","manbin",plrich,div) # columns (response)
  rownames(t)<-colnames(t) # rows (predictor)

  for (k in rownames(t)) {
    for (i in colnames(t)) {
      
      if (i %in% v$coefficients$Predictor[v$coefficients$Response==k]){
      t[i,k]<-v$coefficients$Std.Estimate[v$coefficients$Predictor==i & v$coefficients$Response==k]
      }}}
  
  t[sampltemp,"tempy"]<-v$coefficients$Std.Estimate[v$coefficients$Predictor==paste0("~~",sampltemp) & v$coefficients$Response=="~~tempy"]
  t["tempy",sampltemp]<-t[sampltemp,"tempy"]
  
  return(t)
}

#### Variable positions ####

# this matrix lists where in a 3x4 matrix the column names from table t should be placed
l<-matrix(c(rep(0,12)),4,3)

l[1,1]=1
l[2,1]=2
l[3,1]=3
l[4,1]=4
l[3,3]=5
l[1,3]=6

#### Arrow labels ####

# function to transform p values into asterisks
psign<-function(p){
if(p<=0.001) {
  p <- "***"
  } else if (p<=0.01) {
    p <- "**"
    } else if (p<=0.05) {
      p <- "*"
      } else if (p<=0.1) {
        p <- "."
        } else {p <- "n.s."}
  p
}

# order according to number of response variable (that's why first k, then i)

edgelabels<-function(v, sampltemp, div) {
  n<-3; sl<-c(NA, NA)

  for (k in c(sampltemp,"tempy","pp","manbin","pldiv","flrich", div)) { # columns (response)
    for (i in c(sampltemp,"tempy","pp","manbin","pldiv","flrich", div)) { # rows (predictor)
      if (i %in% v$coefficients$Predictor[v$coefficients$Response==k]) {
        sl[n]<-paste(round(v$coefficients$Std.Estimate[v$coefficients$Predictor==i & v$coefficients$Response==k], 2),psign(v$coefficients$P.Value[v$coefficients$Predictor==i & v$coefficients$Response==k]))
        n<-n+1
      }}}
  return(sl)
}

#### Arrow colors ####

# function to transform p values into colour codes
pcolpos<-function(p){
if(p<=0.001) {
  p <- "darkgreen"
  } else if (p<=0.01) {
    p <- "darkgreen"
    } else if (p<=0.05) {
      p <- "darkgreen"
      } else if (p<=0.1) {
        p <- "slategray2"
        } else {p <- "slategray2"}
  p
}

pcolneg<-function(p){
if(p<=0.001) {
  p <- "red4"
  } else if (p<=0.01) {
    p <- "red4"
    } else if (p<=0.05) {
      p <- "red4"
      } else if (p<=0.1) {
        p <- "slategray2"
        } else {p <- "slategray2"}
  p
}


edgecolors<-function(v, sampltemp, div) {
  n<-3; sl<-c("black", "black")
  
for (k in c(sampltemp,"tempy","pp","manbin","pldiv","flrich", div)) { # columns (response)
    for (i in c(sampltemp,"tempy","pp","manbin","pldiv","flrich", div)) { # rows (predictor)
      if (i %in% v$coefficients$Predictor[v$coefficients$Response==k]) {
        if (v$coefficients$Estimate[v$coefficients$Predictor==i & v$coefficients$Response==k]>0){
          sl[n]<-pcolpos(v$coefficients$P.Value[v$coefficients$Predictor==i & v$coefficients$Response==k])
        } else if (v$coefficients$Estimate[v$coefficients$Predictor==i & v$coefficients$Response==k]<0){
          sl[n]<-pcolneg(v$coefficients$P.Value[v$coefficients$Predictor==i & v$coefficients$Response==k])
        }
        
        n<-n+1
      }}}
  return(sl)
}
```

### 3. pollinators
#### malaise traps & transect walks
```{r methplots}

# malaise traps
pollmalsm<-summary(psem(pp~1,
                        lm(flrich~manbin+pp+tempy, pd),
                        lm(clpolldiv~flrich+tempsampl, pd),
                        tempy %~~% tempsampl),
                  .progressBar = F)

# transect walks
polldipsm<-summary(psem(pp~1,
                        lm(flrich~manbin+pp+tempy, pd),
                        lm(clpolldip~samplpoll+flrich, pd),
                        tempy %~~% samplpoll),
                   .progressBar = F)

# Estimates
tpollmal<-estimates(pollmalsm, "tempsampl", "clpolldiv", "flrich")
tpolldip<-estimates(polldipsm, "samplpoll", "clpolldip", "flrich")

# Edge labels
elpollmal<-edgelabels(pollmalsm, "tempsampl", "clpolldiv")
elpolldip<-edgelabels(polldipsm, "samplpoll", "clpolldip")

#Edge colors
ecpollmal<-edgecolors(pollmalsm, "tempsampl", "clpolldiv")
ecpolldip<-edgecolors(polldipsm, "samplpoll", "clpolldip")

# plot
plotpaths<-function(div, plrich, sampltemp, method, edgevaluematrix, edgecolor, edgelabel, pathmodel, letter) {

  png(here::here(figures_dir, "sems-single", paste0(div, "sm.png")), height=9, width=8, res=600, units = "cm")

  plot(qgraph(edgevaluematrix, shape="square", layout=l, labels=F, vsize=18, esize=8, edge.color=edgecolor, colFactor=0, asize=6, edge.labels=edgelabel, edge.label.bg=c(NA, NA, rep("white", length(edgelabel)-2)), edge.label.cex=1.8, edge.label.margin=0.025, edge.label.position=0.7, mar=c(3,4,3,3), images=c( # edge.label.position normal 0.5
    here::here(figures_dir, "logos", paste0(sampltemp, ".png")),
    here::here(figures_dir, "logos", "tempy.png"),
    here::here(figures_dir, "logos", "pp.png"),
    here::here(figures_dir, "logos", "manbin.png"),
    here::here(figures_dir, "logos", paste0(plrich, ".png")),
    here::here(figures_dir, "logos", paste0(div, ".png"))), 
    borders = F # no border lines around logos
        ) %>% inset2("Edgelist", (.) %>% extract2("Edgelist") %>% inset2("bidirectional", c(TRUE,TRUE, rep(FALSE,length(edgelabel)-2)))) # correct temperature to double arrow
        ); grid.raster(readPNG(here::here(figures_dir, "logos", paste0(method, ".png"))), width = .17, x=.9, y=.78 # add method logo, coordinates before x .75, y .78
        ); text(1,1.25,paste("R²=",pathmodel$R2$R.squared[pathmodel$R2$Response==div]),cex=.8
        ); text(1,-0.65,paste("R²=",pathmodel$R2$R.squared[pathmodel$R2$Response==plrich]),cex=.8
        ); text(-1.25,1.1,letter,cex=1.3
        ); text(0.4,-1.1,paste("Fisher's C =", pathmodel$Cstat$Fisher.C, "\np-value = ", pathmodel$Cstat$P.Value), adj=0, cex=.7)
  
  dev.off()
}

plotpaths("clpolldiv", "flrich", "tempsampl", "malaisetrap", tpollmal, ecpollmal, elpollmal, pollmalsm, "a)")
plotpaths("clpolldip", "flrich", "sampldip", "kescher", tpolldip, ecpolldip, elpolldip, polldipsm, "b)")

pathpollmal<-readPNG(here::here(figures_dir, "sems-single", "clpolldivsm.png"))
pathpolldip<-readPNG(here::here(figures_dir, "sems-single", "clpolldipsm.png"))
```

#### transect walks with abundance

```{r abplot}
pollabsm<-summary(psem(pp~1,
                      lm(flrich~manbin+pp+tempy, pd),
                      lm(clpollab~samplpoll+flrich, pd), 
                      lm(clpolldip~clpollab+flrich, pd),
                      tempy %~~% samplpoll),
                  .progressBar = F)

#### Estimates ####

# manual matrix with estimates from path model for plotting

estimates2<-function(v, sampltemp, div, ab, plrich) {
  
  t<-matrix(c(rep(0,7*7)),7,7,byrow=T)

  colnames(t)<-c(sampltemp,"tempy","pp","manbin",plrich,div,ab) # columns (response)
  rownames(t)<-colnames(t) # rows (predictor)

  for (k in rownames(t)) {
    for (i in colnames(t)) {
      
      if (i %in% v$coefficients$Predictor[v$coefficients$Response==k]){
      t[i,k]<-v$coefficients$Std.Estimate[v$coefficients$Predictor==i & v$coefficients$Response==k]
      }}}
  
  t[sampltemp,"tempy"]<-v$coefficients$Std.Estimate[v$coefficients$Predictor==paste0("~~",sampltemp) & v$coefficients$Response=="~~tempy"]
  t["tempy",sampltemp]<-t[sampltemp,"tempy"]
  
  return(t)
}

tpollab<-estimates2(pollabsm, "samplpoll", "clpolldip", "clpollab", "flrich")

#### Variable positions ####

# this matrix lists where in a 3x4 matrix the column names from table t should be placed
l2<-matrix(c(rep(0,24)),8,3)

l2[1,1]=1
l2[3,1]=2
l2[5,1]=3
l2[7,1]=4
l2[5,3]=5
l2[1,3]=6
l2[2,2]=7

#### Arrow labels ####

# order according to number of response variable (that's why first k, then i)

edgelabels2<-function(v, sampltemp, div, ab) {
  n<-3; sl<-c(NA, NA)

  for (k in c(sampltemp,"tempy","pp","manbin","pldiv","flrich",div,ab)) { # columns (response)
    for (i in c(sampltemp,"tempy","pp","manbin","pldiv","flrich",div,ab)) { # rows (predictor)
      if (i %in% v$coefficients$Predictor[v$coefficients$Response==k]) {
        sl[n]<-paste(round(v$coefficients$Std.Estimate[v$coefficients$Predictor==i & v$coefficients$Response==k], 2),psign(v$coefficients$P.Value[v$coefficients$Predictor==i & v$coefficients$Response==k]))
        n<-n+1
      }}}
  return(sl)
}

elpollab<-edgelabels2(pollabsm, "samplpoll", "clpolldip", "clpollab")

#### Arrow colors ####

edgecolors2<-function(v, sampltemp, div, ab) {
  n<-3; sl<-c("black", "black")
  
for (k in c(sampltemp,"tempy","pp","manbin","pldiv","flrich",div,ab)) { # columns (response)
    for (i in c(sampltemp,"tempy","pp","manbin","pldiv","flrich",div,ab)) { # rows (predictor)
      if (i %in% v$coefficients$Predictor[v$coefficients$Response==k]) {
        if (v$coefficients$Estimate[v$coefficients$Predictor==i & v$coefficients$Response==k]>0){
          sl[n]<-pcolpos(v$coefficients$P.Value[v$coefficients$Predictor==i & v$coefficients$Response==k])
        } else if (v$coefficients$Estimate[v$coefficients$Predictor==i & v$coefficients$Response==k]<0){
          sl[n]<-pcolneg(v$coefficients$P.Value[v$coefficients$Predictor==i & v$coefficients$Response==k])
        }
        
        n<-n+1
      }}}
  return(sl)
}

ecpollab<-edgecolors2(pollabsm, "samplpoll", "clpolldip", "clpollab")

#### plot ####

plotpaths2<-function(div, ab, plrich, sampltemp, edgevaluematrix, edgecolor, edgelabel, pathmodel, letter) {
  
  svg(here::here(figures_dir, "sems-single", paste0(div, "absm.svg")), height=3.54, width=3.15)
  
  plot(qgraph(edgevaluematrix, shape="square", layout=l2, labels=F, vsize=18, esize=8, edge.color=edgecolor, colFactor=0, asize=6, edge.labels=edgelabel, edge.label.bg=c(NA, NA, rep("white", length(edgelabel)-2)), edge.label.cex=1.8, edge.label.margin=0.025, edge.label.position=0.65, mar=c(3,4,3,3), images=c(
    here::here(figures_dir, "logos", paste0(sampltemp, ".png")),
    here::here(figures_dir, "logos", "tempy.png"),
    here::here(figures_dir, "logos", "pp.png"),
    here::here(figures_dir, "logos", "manbin.png"),
    here::here(figures_dir, "logos", paste0(plrich, ".png")),
    here::here(figures_dir, "logos", paste0(div, ".png")),
    here::here(figures_dir, "logos", paste0(ab, ".png"))), 
    borders = F # no border lines around logos
        ) %>% inset2("Edgelist", (.) %>% extract2("Edgelist") %>% inset2("bidirectional", c(TRUE,TRUE, rep(FALSE,length(edgelabel)-2))))); # correct temperature to double arrow
  text(1,1.25,paste("R²=",pathmodel$R2$R.squared[pathmodel$R2$Response==div]),cex=.8
        );text(1,-0.65,paste("R²=",pathmodel$R2$R.squared[pathmodel$R2$Response==plrich]),cex=.8
        );text(0,.97,paste("R²=",pathmodel$R2$R.squared[pathmodel$R2$Response==ab]),cex=.8
        );text(-1.25,1.1,letter,cex=1.3
        );text(0.4,-1.1,paste("Fisher's C =", pathmodel$Cstat$Fisher.C, "\np-value = ", pathmodel$Cstat$P.Value), adj=0, cex=.7)
  
  dev.off()
}

plotpaths2("clpolldip", "clpollab", "flrich", "sampldip", tpollab, ecpollab, elpollab, pollabsm, "c)") # exported as svg to manually adjust arrow labels & save it as png

pathpollab<-readPNG(here::here(figures_dir, "sems-single", "clpolldipabsm.png")) # read manually adjusted and saved png file
```

#### legend

```{r polllegend}
tlegend<-matrix(c(rep(0,16*16)),16,16,byrow=T)
colnames(tlegend)<-c("tempsampl", "sampling temperature malaise", "tempy", "annual temperature", "pp", "primary productivity", "manbin", "management", "polldiv", "sampldip", "sampling temperature transects", "pollinator richness", "clpollab", "pollinator abundance", "flrich", "flower richness") # Spalten (response)
rownames(tlegend)<-colnames(tlegend)

ll<-matrix(c(rep(0,16)),4,4)

ll[1,1]=1; ll[1,2]=2
ll[2,1]=3; ll[2,2]=4
ll[3,1]=5; ll[3,2]=6
ll[4,1]=7; ll[4,2]=8
ll[1,3]=9; ll[1,4]=10
ll[2,3]=11; ll[2,4]=12
ll[3,3]=13; ll[3,4]=14
ll[4,3]=15; ll[4,4]=16

png(here::here(figures_dir, "sems-single", "legendpoll.png"), height=9, width=8, res=600, units = "cm")
  
qgraph(tlegend, shape="square", layout=ll, labels=F, vsize=18, images=c(
  here::here(figures_dir, "logos", "tempsampl.png"), NA,
  here::here(figures_dir, "logos", "tempy.png"), NA,
  here::here(figures_dir, "logos", "pp.png"), NA,
  here::here(figures_dir, "logos", "manbin.png"), NA,
  here::here(figures_dir, "logos", "sampldip.png"), NA, 
  here::here(figures_dir, "logos", "polldiv.png"), NA, 
  here::here(figures_dir, "logos", "clpollab.png"), NA, 
  here::here(figures_dir, "logos", "flrich.png"), NA),
  borders = F # no border lines around logos
    );text(-0.7,1,"sampling\ntemperature\nmalaise trap",cex=.9, adj = 0
    );text(-0.7,0.35,"annual\ntemperature",cex=.9, adj = 0
    );text(-0.7,-0.35,"primary\nproductivity",cex=.9, adj = 0
    );text(-0.7,-1,"management",cex=.9, adj = 0
    );text(0.63,1,"sampling\ntemperature\ntransects",cex=.9, adj = 0
    );text(0.63,0.35,"pollinator\nrichness",cex=.9, adj = 0
    );text(0.63,-0.35,"pollinator\nabundance",cex=.9, adj = 0
    );text(0.63,-1,"flower\nrichness",cex=.9, adj = 0)
     
dev.off()

pathlegendpoll<-readPNG(here::here(figures_dir, "sems-single", "legendpoll.png"))
```

#### assemble plots

```{r pollsemplot, fig.cap="", echo=FALSE, warning=FALSE, message=FALSE, error=FALSE, fig.width=16, fig.height=11}
png(here::here(figures_dir, "sempoll.png"), height=11, width=10, res=600, units = "cm")

grid.arrange(rasterGrob(pathpollmal), rasterGrob(pathlegendpoll), rasterGrob(pathpolldip), rasterGrob(pathpollab), ncol=2)

dev.off()
```

### 4. grasshoppers
#### malaise traps & transect walks

```{r orpathmodels}
# malaise traps
ormalsm<-summary(psem(pp~1,
                      lm(pldiv~manbin+pp+tempy, pd),
                      lm(divor~manbin+pp+tempy, pd),
                      tempy %~~% tempsampl),
                 .progressBar = F)

# transect walks
ordipsm<-summary(psem(pp~1,
                      lm(pldiv~manbin+pp+tempy, pd),
                      lm(ordip~tempy, pd),
                      tempy %~~% samplor),
                 .progressBar = F)

# Estimates
tmalor<-estimates(ormalsm, "tempsampl", "divor", "pldiv")
tdipor<-estimates(ordipsm, "samplor", "ordip", "pldiv")

# Edge labels
elmalor<-edgelabels(ormalsm, "tempsampl", "divor")
eldipor<-edgelabels(ordipsm, "samplor", "ordip")

# Edge colors
ecmalor<-edgecolors(ormalsm, "tempsampl", "divor")
ecdipor<-edgecolors(ordipsm, "samplor", "ordip")

# plot
plotpaths("divor", "pldiv", "tempsampl", "malaisetrap", tmalor, ecmalor, elmalor, ormalsm, "a)")
plotpaths("ordip", "pldiv", "sampldip", "kescher", tdipor, ecdipor, eldipor, ordipsm, "b)")

ormal<-readPNG(here::here(figures_dir, "sems-single", "divorsm.png"))
ordip<-readPNG(here::here(figures_dir, "sems-single", "ordipsm.png"))
```

#### legend

```{r polllegend}
# legend
tlegend2<-matrix(c(rep(0,14*14)),14,14,byrow=T)
colnames(tlegend2)<-c("tempsampl", "sampling temperature malaise", "tempy", "annual temperature", "pp", "primary productivity", "manbin", "management", "sampldip", "sampling temperature transects", "divor", "grasshopper richness", "pldiv", "plant richness") # columns(response)
rownames(tlegend2)<-colnames(tlegend2)

ll2<-matrix(c(rep(0,16)),4,4)

ll2[1,1]=1; ll2[1,2]=2
ll2[2,1]=3; ll2[2,2]=4
ll2[3,1]=5; ll2[3,2]=6
ll2[4,1]=7; ll2[4,2]=8
ll2[1,3]=9; ll2[1,4]=10
ll2[2,3]=11; ll2[2,4]=12
ll2[3,3]=13; ll2[3,4]=14
ll2[4,3]=15; ll2[4,4]=16

png(here::here(figures_dir, "sems-single", "legendor.png"), height=9, width=8, res=600, units = "cm")
  
qgraph(tlegend2, shape="square", layout=ll2, labels=F, vsize=18, images=c(
  here::here(figures_dir, "logos", "tempsampl.png"), NA,
  here::here(figures_dir, "logos", "tempy.png"), NA,
  here::here(figures_dir, "logos", "pp.png"), NA,
  here::here(figures_dir, "logos", "manbin.png"), NA,
  here::here(figures_dir, "logos", "sampldip.png"), NA, 
  here::here(figures_dir, "logos", "divor.png"), NA, 
  here::here(figures_dir, "logos", "pldiv.png"), NA),
  borders = F # no border lines around logos
    );text(-0.7,1,"sampling\ntemperature\nmalaise trap",cex=.9, adj = 0
    );text(-0.7,0.35,"annual\ntemperature",cex=.9, adj = 0
    );text(-0.7,-0.35,"primary\nproductivity",cex=.9, adj = 0
    );text(-0.7,-1,"management",cex=.9, adj = 0
    );text(0.63,1,"sampling\ntemperature\ntransects",cex=.9, adj = 0
    );text(0.63,0.35,"grass-\nhopper\nrichness",cex=.9, adj = 0
    );text(0.63,-0.35,"plant\nrichness",cex=.9, adj = 0)
     
dev.off()

pathlegendor<-readPNG(here::here(figures_dir, "sems-single", "legendor.png"))
```

#### assemble plots

```{r methsemplot, fig.cap="", echo=FALSE, warning=FALSE, message=FALSE, error=FALSE, fig.width=16, fig.height=11}
png(here::here("text", "figures", "orthmeth.png"), height=16.5, width=5, res=600, units = "cm")

grid.arrange(rasterGrob(ormal), rasterGrob(ordip), rasterGrob(pathlegendor), ncol=1)

dev.off()
```

# Supplementary Material

## S2. comparison species richness of all & selected pollinators

```{r pollpatterns}
pollnames<-c(`polldiv` = "all", `clpolldiv` = "selected")

pollplot<-function(dataframe) {
  dataframe %>% pivot_longer(cols=c(polldiv, clpolldiv),names_to = c("eco.group"),values_to="spec.no") %>%
  ggplot(aes(elev, spec.no, col=eco.group, fill=eco.group)) +
  facet_wrap(vars(fct_inorder(eco.group)),ncol = 2, scales = "free", labeller = as_labeller(pollnames)) +
  geom_point(size=1.5, alpha=0.4, shape=16) +
  geom_smooth(method="gam", formula = y ~ s(x, k=4), se = T, lwd = 1, alpha=0.4) +
  theme_classic() +
  theme(axis.text = element_text(size=9),
        axis.title = element_text(size=10),
        legend.title = element_blank(),
        legend.position = "none",
        strip.text = element_text(size=10), # modify facet label text
        strip.background = element_blank()) + # remove frame from facet labels
  labs(x="elevation [m a.s.l.]") +
  scale_fill_manual(
    values = c("polldiv"="#C73965", "clpolldiv"="#FFB601"),
    labels = c("polldiv"="all pollinators", "clpolldiv"="monitored pollinators"), name="richness") +
  scale_color_manual(
    values = c("polldiv"="#C73965", "clpolldiv"="#FFB601"),
    labels = c("polldiv"="all pollinators", "clpolldiv"="monitored pollinators"), name="richness") +
    ylab("species richness")
}

pollcorval<-cor.test(pd$polldiv, pd$clpolldiv, method="pearson")

pollcor<-ggplot(aes(polldiv, clpolldiv), data=pd) +
  geom_point(pch=16, alpha=0.5, col="#F79646") +
  geom_smooth(formula = y ~ x, se=F, method="lm", col="#F79646") +
  theme_classic() +
  # ylim(0,17) +
  annotate("text", label = paste("cor =", round(pollcorval$estimate, 2)),
    x = 170, y = 60) +
  labs(x="all pollinators", y="selected pollinators")


png(here::here(suppl_dir, "pollcorr.png"), height=13.5, width=9, res=600, units = "cm")

pollplot(pd) / pollcor + plot_layout(heights = c(1,1.8))

dev.off()
```

## S3. correlation elevation with annual & sampling temperatures

```{r pldivelevcor}
tempcheck<-pivot_longer(data=pd, cols=c(tempy, tempsampl, samplpoll),names_to = c("group"),values_to="temp") %>%
  ggplot(aes(elev, temp)) +
  facet_wrap(vars(group),ncol = 1, scales = "free", labeller = as_labeller(c("tempy" = "annual", "samplpoll"="transect walks", "tempsampl"="malaise traps"))) +
  geom_point(pch=16, alpha=0.5, col="coral2") +
  theme_classic() +
  labs(x="elevation [m a.s.l.]", y="temperature [°C]")

png(here::here(suppl_dir, "temppatterns.png"), height=15, width=6, res=600, units = "cm")

tempcheck

dev.off()
```

## S4. pollinator path model hypothesized causal structure

```{r basepoll}
# basic model
tbase<-matrix(c(rep(0,6*6)),6,6,byrow=T)
colnames(tbase)<-c("tempsampl","tempy","pp","manbin","pldiv","div") # columns (response)
rownames(tbase)<-colnames(tbase)
tbase["tempy","tempsampl"]=0.1
tbase["tempsampl","tempy"]=0.1
tbase["tempsampl","pp"]=0.1
tbase["tempy","pp"]=0.1
tbase["tempy","pldiv"]=0.1
tbase["pp","pldiv"]=0.1
tbase["manbin","pldiv"]=0.1
tbase["tempsampl","div"]=0.1
tbase["tempy","div"]=0.1
tbase["pp","div"]=0.1
tbase["manbin","div"]=0.1
tbase["pldiv","div"]=0.1

png(here::here(figures_dir, "sems-single", "basesmmeth.png"), height=9, width=8, res=600, units = "cm")

plot(qgraph(tbase, shape="square", layout=l, labels=F, vsize=18, edge.color="black", asize=6, edge.labels=F, curveAll=T, curve=c(rep(0,2),-2.5,rep(0,9)), mar=c(3,4,3,3), images=c( # curve=how bend the curve will be, curveAll to not just bend when there are several arrows on a variable
    here::here(figures_dir, "logos", "tempsampl.png"),
    here::here(figures_dir, "logos", "tempy.png"),
    here::here(figures_dir, "logos", "pp.png"),
    here::here(figures_dir, "logos", "manbin.png"),
    here::here(figures_dir, "logos", "flrich.png"),
    here::here(figures_dir, "logos", "polldiv.png")), 
    borders = F # no border lines around logos
      ) %>% inset2("Edgelist", (.) %>% extract2("Edgelist") %>% inset2("bidirectional", c(TRUE,TRUE, rep(FALSE,10))))); text(-1.25,1.1,"",cex=1.3)

dev.off()

pathbasemeth<-readPNG(here::here(figures_dir, "sems-single", "basesmmeth.png"))
```

```{r polllegend}
# legend
tlegend3<-matrix(c(rep(0,16*16)),16,16,byrow=T)
colnames(tlegend3)<-c("tempsampl", "sampling temperature malaise", "tempy", "annual temperature", "pp", "primary productivity", "manbin", "management", "polldiv", "pollinator richness", "flrich", "flower richness", NA, NA, NA, NA) # columns (response)
rownames(tlegend3)<-colnames(tlegend3)

ll3<-matrix(c(rep(0,16)),4,4)

ll3[1,1]=1; ll3[1,2]=2
ll3[2,1]=3; ll3[2,2]=4
ll3[3,1]=5; ll3[3,2]=6
ll3[4,1]=7; ll3[4,2]=8
ll3[1,3]=9; ll3[1,4]=10
ll3[2,3]=11; ll3[2,4]=12
ll3[3,3]=13; ll3[3,4]=14
ll3[4,3]=15; ll3[4,4]=16

png(here::here(figures_dir, "sems-single", "legendbasepoll.png"), height=9, width=8, res=600, units = "cm")
  
qgraph(tlegend3, shape="square", layout=ll3, labels=F, vsize=18, images=c(
  here::here(figures_dir, "logos", "tempsampl.png"), NA,
  here::here(figures_dir, "logos", "tempy.png"), NA,
  here::here(figures_dir, "logos", "pp.png"), NA,
  here::here(figures_dir, "logos", "manbin.png"), NA,
  here::here(figures_dir, "logos", "polldiv.png"), NA, 
  here::here(figures_dir, "logos", "flrich.png"), NA, NA, NA, NA, NA),
  borders = F # no border lines around logos
    );text(-0.7,1,"sampling\ntemperature",cex=.9, adj = 0
    );text(-0.7,0.35,"annual\ntemperature",cex=.9, adj = 0
    );text(-0.7,-0.35,"primary\nproductivity",cex=.9, adj = 0
    );text(-0.7,-1,"management",cex=.9, adj = 0
    );text(0.63,1,"pollinator\nrichness",cex=.9, adj = 0
    );text(0.63,0.35,"flower\nrichness",cex=.9, adj = 0
    )
     
dev.off()

pathlegendbasepoll<-readPNG(here::here(figures_dir, "sems-single", "legendbasepoll.png"))
```

```{r pollbaseplot}
png(here::here(suppl_dir, "basesempoll.png"), height=5.5, width=10, res=600, units = "cm")

grid.arrange(rasterGrob(pathbasemeth), rasterGrob(pathlegendbasepoll), ncol=2)

dev.off()
```

## S5. Venn diagrams individual species groups

```{r venndiagram}
# function see 1. Venn diagrams - grasshoppers
venny("Diptera", "malaise traps", "transect walks", 200, 160, 0, 60, -0.2, 0, 0.055, 1.1)
venny("Hymenoptera", "malaise traps", "transect walks", 20, 340, 180, 60, -0.2, 0, 0.055, 1.1)
venny("Lepidoptera", "malaise traps", "transect walks", 20, 340, 180, 355, -0.03, 0.75, 0.055, 1.1)

vendip<-readPNG(here::here(figures_dir, "venns-single", "venndiagDiptera.png"))
venhym<-readPNG(here::here(figures_dir, "venns-single", "venndiagHymenoptera.png"))
venlep<-readPNG(here::here(figures_dir, "venns-single", "venndiagLepidoptera.png"))

png(here::here("text", "supplementary", "vennspec.png"), height=3.5, width=10.5, res=600, units = "cm")

grid.arrange(rasterGrob(vendip), rasterGrob(venhym), rasterGrob(venlep), nrow=1);grid.raster(
  readPNG(here::here(figures_dir, "logos", "hvdip.png")), width = .055, x=.035, y=.91);grid.raster(
  readPNG(here::here(figures_dir, "logos", "wbdip.png")), width = .055, x=.365, y=.91);grid.raster(
  readPNG(here::here(figures_dir, "logos", "bfdip.png")), width = .06, x=.695, y=.91)

dev.off()
```

## S6. body size comparisons

```{r body sizes}
MT<-read.csv(here::here(data_proc, "body_sizes.csv"), row.names = 1)

library(ggtext) # including images for x axis labels

# define images as labels (html code included)

imgpaths<-c(here::here(figures_dir, "logos", "divwb.png"), here::here(figures_dir, "logos", "divhv.png"),
            here::here(figures_dir, "logos", "divbf.png"), here::here(figures_dir, "logos", "divor.png"))

labels <- c()

for (i in 1:length(levels(MT$group))){
  
  img.name <- levels(MT$group)[i]
  
  labels <- c(labels, paste0("<img src='", imgpaths[i],  "' width='25' /><br>*", img.name,"*"))
  
}

png(here::here("text", "supplementary", "body_sizes_methods.png"), height=9, width=13.5, res=600, units = "cm")

ggplot(MT, aes(col=method1, fill=method1, y=bodysize, x=group, alpha=method1)) + 
  geom_boxplot(outlier.shape=16, outlier.size=2, notch=FALSE) +
  theme_classic() +
  scale_color_manual(values = c("malaise traps"="#21908dff", "transect walks"="#440154ff"),
                    name=element_blank()) +
  scale_fill_manual(values = c("malaise traps"="#21908dff", "transect walks"="#440154ff"),
                    name=element_blank()) +
  scale_alpha_manual(values = c("malaise traps"=.5, "transect walks"=.5), name=element_blank()) +
  labs(x=element_blank(), y="body size [mm]") +
  scale_x_discrete(name = NULL, labels = labels) + # use images for axis labels
  theme(axis.text.x = element_markdown(), # add text to image labels
    legend.position=c(.14, .93))

dev.off()
```

## S7. beta diversity

```{r beta}

png(here::here("text", "supplementary", "betadiversity.png"), height=8, width=11, res=600, units = "cm")

pd %>% pivot_longer(cols=c(betadivj, betadipj),names_to = c("method"),values_to="beta") %>%
  ggplot(aes(elev, beta, col=method, fill=method)) +
  geom_point(size=1.5, alpha=0.6, shape=16) +
  geom_smooth(aes(linetype=method), method="gam", formula = y ~ s(x, k=4), se = T, lwd = .8) +
  theme_classic() +
  labs(x="elevation [m a.s.l.]", y="beta diversity") +
  theme(legend.position=c(.86, .93)) +
  scale_color_manual(values = c("betadivj"="#21908dff", "betadipj"="#440154ff"),
         labels = c("betadivj"="malaise traps", "betadipj"="transect walks"),
         name=element_blank()) +
  scale_fill_manual(values = c("betadivj"="#21908dff", "betadipj"="#440154ff"),
         labels = c("betadivj"="malaise traps", "betadipj"="transect walks"),
         name=element_blank()) +
  scale_linetype_manual(values = c("betadivj"="solid", "betadipj"="dotdash"),
         labels = c("betadivj"="malaise traps", "betadipj"="transect walks"),
         name=element_blank())

dev.off()
```

## S8 & S9. Elevational & sampling temperature patterns for individual pollinator groups

```{r methspecgroupmodelline}
# read analysis
rmarkdown::render(here::here("results", "scripts", "Models.Rmd"), quiet = T) # quiet to not get text from rendering

# Labels
elevtags<-data.frame(lab=c(paste(psign(elevhv$p.pv["method.L"]),"/",psign(elevhv$s.table[2,4])),
                           paste(psign(elevwb$p.pv["method.L"]),"/",psign(elevwb$s.table[2,4])),
                           paste(psign(elevbf$p.pv["method.L"]),"/",psign(elevbf$s.table[2,4]))),
                     group=c("hoverflies","wild bees", "butterflies"))

temptags<-data.frame(lab=c(paste(psign(temphv$p.pv["method.L"]),"/",psign(temphv$s.table[2,4])),
                           paste(psign(tempwb$p.pv["method.L"]),"/",psign(tempwb$s.table[2,4])),
                           paste(psign(tempbf$p.pv["method.L"]),"/",psign(tempbf$s.table[2,4]))),
                     group=c("hoverflies","wild bees", "butterflies"))

elevr2<-data.frame(lab=c(paste("R² ≈", round(elevhv$r.sq, 2)),
                         paste("R² ≈", round(elevwb$r.sq, 2)),
                         paste("R² ≈", round(elevbf$r.sq, 2))),
                   group=c("hoverflies","wild bees", "butterflies"))

tempr2<-data.frame(lab=c(paste("R² ≈", round(temphv$r.sq, 2)),
                         paste("R² ≈", round(tempwb$r.sq, 2)),
                         paste("R² ≈", round(tempbf$r.sq, 2))),
                   group=c("hoverflies","wild bees", "butterflies"))

# syrphids
elevhv<-gam(spec.no~s(elev, k=4)+s(elev,by=method,k=4)+method+s(plot, bs="re"), data=sm[sm$group=="hoverflies",], family=poisson, method="REML")

elevhv<-get_gam_predictions(elevhv, elev, series_length = 10*(max(sm$elev)-min(sm$elev))+1)
elevhv$group<-"hoverflies"

temphv<-gam(spec.no~s(tempright, k=4)+s(tempright,by=method,k=4)+method+s(elev, k=4)+s(plot, bs="re"), data=sm[sm$group=="hoverflies",], family=poisson, method="REML")

temphv<-get_gam_predictions(temphv, tempright, series_length = 100*(max(sm$tempright)-min(sm$tempright))+1)
temphv$group<-"hoverflies"

# wild bees
elevwb<-gam(spec.no~s(elev, k=4)+s(elev,by=method,k=4)+method+s(plot, bs="re"), data=sm[sm$group=="wild bees",], family=poisson, method="REML")

elevwb<-get_gam_predictions(elevwb, elev, series_length = 10*(max(sm$elev)-min(sm$elev))+1)
elevwb$group<-"wild bees"

tempwb<-gam(spec.no~s(tempright, k=4)+s(tempright,by=method,k=4)+method+s(elev, k=4)+s(plot, bs="re"), data=sm[sm$group=="wild bees",], family=poisson, method="REML")

tempwb<-get_gam_predictions(tempwb, tempright, series_length = 100*(max(sm$tempright)-min(sm$tempright))+1)
tempwb$group<-"wild bees"

# butterflies
elevbf<-gam(spec.no~s(elev, k=4)+s(elev,by=method,k=4)+method+s(plot, bs="re"), data=sm[sm$group=="butterflies",], family=poisson, method="REML")

elevbf<-get_gam_predictions(elevbf, elev, series_length = 10*(max(sm$elev)-min(sm$elev))+1)
elevbf$group<-"butterflies"

tempbf<-gam(spec.no~s(tempright, k=4)+s(tempright,by=method,k=4)+method+s(elev, k=4)+s(plot, bs="re"), data=sm[sm$group=="butterflies",], family=poisson, method="REML")

tempbf<-get_gam_predictions(tempbf, tempright, series_length = 100*(max(sm$tempright)-min(sm$tempright))+1)
tempbf$group<-"butterflies"

# elev all
allelev<-rbind(elevhv, elevwb, elevbf)

for (i in c("spec.no","SE","CI_upper","CI_lower")){ # backtransform poisson (log)
  allelev[i]<-exp(allelev[i])
}

# temp all
alltemp<-rbind(temphv[temphv$method=="malaise traps" &
                             temphv$tempright>=min(sm$tempsampl) &
                             temphv$tempright<=max(sm$tempsampl),],
               temphv[temphv$method=="transect walks" &
                             temphv$tempright>=min(sm$samplhv) &
                             temphv$tempright<=max(sm$samplhv),],
               tempwb[tempwb$method=="malaise traps" &
                             tempwb$tempright>=min(sm$tempsampl) &
                             tempwb$tempright<=max(sm$tempsampl),],
               tempwb[tempwb$method=="transect walks" &
                             tempwb$tempright>=min(sm$samplwb) &
                             tempwb$tempright<=max(sm$samplwb),],
               tempbf[tempbf$method=="malaise traps" &
                             tempbf$tempright>=min(sm$tempsampl) &
                             tempbf$tempright<=max(sm$tempsampl),],
               tempbf[tempbf$method=="transect walks" &
                             tempbf$tempright>=min(sm$samplbf) &
                             tempbf$tempright<=max(sm$samplbf),])

for (i in c("spec.no","SE","CI_upper","CI_lower")){ # backtransform poisson (log)
  alltemp[i]<-exp(alltemp[i])
}

# plot
methspecplot<-function (var, xaxisname, xcoord, ycoord, tags, tagpos.x, tagpos.y, xcorrect, ycorrect, dataframe, r2, rpos.x, rpos.y, rxcorrect, rycorrect) {
  g <- ggplot_gtable(ggplot_build(
    ggplot(aes({{var}}, spec.no, col=method), data=dataframe) +
    facet_wrap(vars(fct_inorder(group)), ncol = 1, scales = "free") +
    geom_ribbon(aes(ymin = CI_lower, ymax = CI_upper, fill=method), alpha=.4, colour = NA) +
    geom_point(pch=16, alpha=0.6, data=sm) +
    geom_line(aes(linetype=method, col=method), lwd=0.7) +
    theme_classic() +
    theme(axis.text = element_text(size=10),
          axis.title = element_text(size=10),
          strip.text = element_blank(), # suppress plot labelling by facetwrap
          legend.position = "none") +
    labs(x=xaxisname, y="species richness") +
    scale_color_manual(values = c("malaise traps"="#21908dff", "transect walks"="#440154ff"),
    name="sampling\nmethod") +
    scale_fill_manual(values = c("malaise traps"="#21908dff", "transect walks"="#440154ff"),name="sampling\nmethod") +
    scale_linetype_manual(values = c("malaise traps"="solid", "transect walks"="dotdash"),
      name="sampling\nmethod") +
    scale_x_continuous(n.breaks=4) +
    coord_cartesian(ylim = c(0, NA)) + # set lower limit to 0 without stopping to plot se
    geom_text(inherit.aes = FALSE, data=tags, aes(x=tagpos.x, y=tagpos.y, label=lab), col="black", hjust=xcorrect, vjust=ycorrect, size=3.5) +
    geom_text(inherit.aes = FALSE, data=r2, aes(x=rpos.x, y=rpos.y, label=lab), col="black", hjust=rxcorrect, vjust=rycorrect, size=3.5)
  ))

  # add logos
  
  facets <- grep("panel", g$layout$name)
  
  logos <- list(rasterGrob(readPNG(here::here(figures_dir, "logos", "hvdip.png")),
                             width=.15, x = xcoord, y = ycoord),
                  rasterGrob(readPNG(here::here(figures_dir, "logos", "wbdip.png")),
                             width=.15, x = xcoord, y = ycoord),
                  rasterGrob(readPNG(here::here(figures_dir, "logos", "bfdip.png")),
                             width=.15, x = xcoord, y = .9))
  
  g2 <- with(g$layout[facets,],
          gtable::gtable_add_grob(g, logos,
                          t=t, l=l, b=b, r=r))
  
  plot(g2)
  }

# elevation
png(here::here("text", "supplementary", "elevpatterns-specgroups.png"), height=15, width=6, res=600, units = "cm")

methspecplot(elev, "elevation [m a.s.l.]", .9, .88, elevtags, 1800, Inf, 1, 1.1, allelev, elevr2, -Inf, -Inf, -.15, -.8)

dev.off()

# sampling temperature
png(here::here("text", "supplementary", "temppatterns-specgroups.png"), height=15, width=6, res=600, units = "cm")

methspecplot(tempright, "sampling temperature [°C]", .1, .88, temptags, -Inf, Inf, -1, 1.2, alltemp, tempr2, Inf, -Inf, 1.1, -.8)

dev.off()
```
