---
title: "Models"
output: html_document
---

load libraries

```{r libraries, include=FALSE}
library(lmerTest)
library(mgcv) # GAMs
library(nlme) # linear models with random term(s)
library(piecewiseSEM) # calculate SEMs
library(plyr) # revalue factor levels
library(tidymv) # GAM predictions
library(tidyverse)
```

read data

```{r loaddata, include=FALSE}
pd<-read.csv(here::here("data", "plotdata.csv"), sep=";")

# transformation of not normally distributed annual temperature
pd$tempy<-log(max(pd$tempy+1)-pd$tempy); pd$tempy<-max(pd$tempy)-pd$tempy
```

# Main text

## Elevational & sampling temperature patterns

### pollinators

```{r methodpolldata}
mp<-pivot_longer(data=pd, cols=c(clpolldiv, clpolldip),names_to = c("method"),values_to="spec.no") 

mp$method <-revalue(mp$method, c("clpolldiv"="malaise traps", "clpolldip"="transect walks"))
mp$method<-factor(mp$method, ordered = T) # factor needs to be ordered for ANOVA like GAM summary

# adjusting sample temperature to use only sampling periods of each method
mp$tempright <-case_when(mp$method =="transect walks" ~ mp$samplpoll,
                         mp$method =="malaise traps" ~ mp$tempsampl)

mp$plot<-as.factor(mp$plot)

# Elevational pattern both methods
pollelev<-summary(gam(spec.no~s(elev, k=4)+s(elev,by=method,k=4)+method+s(plot, bs="re"), data=mp, family=poisson, method="REML"))

# Sampling temperature pattern both methods
polltemp<-summary(gam(spec.no~s(tempright, k=4)+s(tempright,by=method,k=4)+method+s(elev, k=4)+s(plot, bs="re"), data=mp, family=poisson, method="REML"))

# Sampling temperature for each method separately
pollds<-summary(lm(spec.no~tempright+elev, data=mp[mp$method=="transect walks",]))

pollms<-summary(lm(spec.no~tempright+elev, data=mp[mp$method=="malaise traps",]))
```

#### elevational bands pollinator species numbers

```{r pollrichdiffelev}
# check for significant differences in pollinator species numbers for each elevational section (lower elevations 500-1000 m, mid-elevations 1000-1500, highlands >1500 m)

# T-Tests
t1<-t.test(spec.no~method,data=mp[mp$elev<1000,])
t2<-t.test(spec.no~method,data=mp[mp$elev>=1000 & mp$elev<1500,])
t3<-t.test(spec.no~method,data=mp[mp$elev>=1500,])
```

### grasshoppers

```{r methodordata}
gp<-pivot_longer(data=pd, cols=c(divor, ordip),names_to = c("method"),values_to="spec.no") 

gp$method <-revalue(gp$method, c("divor"="malaise traps", "ordip"="transect walks"))
gp$method<-factor(gp$method, ordered = T) # factor needs to be ordered for ANOVA like GAM summary

# adjusting sample temperature to use only sampling periods of each method
gp$tempright <-case_when(gp$method =="transect walks" ~ gp$samplor,
                         gp$method =="malaise traps" ~ gp$tempsampl)

gp$plot<-as.factor(gp$plot)

# Elevational pattern both methods
elevor<-summary(gam(spec.no~s(elev, k=4)+s(elev,by=method,k=4)+method+s(plot, bs="re"), data=gp, family=poisson, method="REML"))

# Sampling temperature pattern both methods
tempor<-summary(gam(spec.no~s(tempright, k=4)+s(tempright,by=method,k=4)+method+s(elev, k=4)+s(plot, bs="re"), data=gp, family=poisson, method="REML"))
```

## Path models

### malaise trap pollinator path model

```{r pollpathfinal}
# 1 min. model
clpollsm<-summary(psem(pp~1,
                  lm(flrich~manbin+pp+tempy, pd),
                  lm(clpolldiv~flrich+tempsampl, pd),
                  tempy %~~% tempsampl),
                  .progressBar = F)
clpollsm
```

### transect walk pollinator path model

```{r clpolldippathfinal}
# 1 min. model without abundance

clpolldipsm<-summary(psem(pp~1,
                    lm(flrich~manbin+pp+tempy, pd),
                    lm(clpolldip~samplpoll+flrich, pd),
                    tempy %~~% samplpoll),
                  .progressBar = F)
clpolldipsm
```

### transect walk pollinator path model including abundance

```{r clpollabpathfinal}
# with abundance included: 2 min. models, one adding annual temperature (tempy) as an explanatory variable for pollinator abundance, but p > 0.05 for tempy anyway, so decision for model without it

clpollabsm<-summary(psem(pp~1,
                    lm(flrich~manbin+pp+tempy, pd),
                    lm(clpollab~samplpoll+flrich, pd), 
                    lm(clpolldip~clpollab+flrich, pd),
                    tempy %~~% samplpoll),
                  .progressBar = F)
clpollabsm
```

### malaise trap grasshopper path model

```{r ordivpathfinal}
# 1 min. model

ordivsm<-summary(psem(pp~1,
              lm(pldiv~manbin+pp+tempy, pd),
              lm(divor~manbin+pp+tempy, pd),
              tempy %~~% tempsampl),
              .progressBar = F)
ordivsm
```

### transect walk grasshopper path model

```{r ordippathfinal}
# 1 min. model

ordipsm<-summary(psem(pp~1,
              lm(pldiv~manbin+pp+tempy, pd),
              lm(ordip~tempy, pd),
              tempy %~~% samplor),
              .progressBar = F)
ordipsm
```

# Supplementary Material

## S6. body size comparisons

```{r bodysizemeth}
MT<-read.csv(here::here("data", "body_sizes.csv"), sep=";")

body<-summary(lme(bodysize~method, random=~1|group, data=MT[!is.na(MT$bodysize),]))
```

## S7. beta diversity

```{r betadiffelev}
bdiv<-summary(gam(betadivj~s(elev, k=4), data=pd, method="REML"))
bdip<-summary(gam(betadipj~s(elev, k=4), data=pd, method="REML"))

# check for significant differences in beta diversity for each elevational section (lower elevations 500-1000 m, mid-elevations 1000-1500, highlands >1500 m)

b <- pivot_longer(cols=c(betadivj, betadipj),names_to = c("method"),values_to="beta", data=pd)

# T-Tests
b1<-t.test(beta~method,data=b[b$elev<1000,])
b2<-t.test(beta~method,data=b[b$elev>=1000 & b$elev<1500,])
b3<-t.test(beta~method,data=b[b$elev>=1500,])
```

## S8 & S9. Elevational & sampling temperature patterns for individual pollinator groups

```{r methodspec}
# restructure data
sm<-pivot_longer(data=pd, cols=c(divwb, wbdip, divhv, hvdip, divbf, bfdip),names_to = c("spec.group"),values_to="spec.no") 
sm$method <-case_when(sm$spec.group %in% c("divwb", "divhv", "divbf") ~ "malaise traps",
            sm$spec.group %in% c("wbdip", "hvdip", "bfdip") ~ "transect walks") # define method
sm$method<-factor(sm$method, ordered = T) # factor needs to be ordered for ANOVA like GAM summary

sm$group <-case_when(sm$spec.group %in% c("divwb", "wbdip") ~ "wild bees",
                     sm$spec.group %in% c("divhv", "hvdip") ~ "hoverflies",
                     sm$spec.group %in% c("divbf", "bfdip") ~ "butterflies") # define species group
sm$group<-factor(sm$group)

# adjusting sample temperature to use only sampling periods of each method
sm$tempright <-case_when(sm$spec.group =="wbdip" ~ sm$samplwb,
                         sm$spec.group =="bfdip" ~ sm$samplbf,
                         sm$spec.group =="hvdip" ~ sm$samplhv,
                         sm$spec.group %in% c("divwb", "divhv", "divbf") ~ sm$tempsampl)

sm$plot<-as.factor(sm$plot) # plot as factor

# butterflies
elevbf<-summary(gam(spec.no~s(elev, k=4)+s(elev,by=method,k=4)+method+s(plot, bs="re"), data=sm[sm$group=="butterflies",], family=poisson, method="REML"))

tempbf<-summary(gam(spec.no~s(tempright, k=4)+s(tempright,by=method,k=4)+method+s(elev, k=4)+s(plot, bs="re"), data=sm[sm$group=="butterflies",], family=poisson, method="REML"))

# wild bees
elevwb<-summary(gam(spec.no~s(elev, k=4)+s(elev,by=method,k=4)+method+s(plot, bs="re"), data=sm[sm$group=="wild bees",], family=poisson, method="REML"))

tempwb<-summary(gam(spec.no~s(tempright, k=4)+s(tempright,by=method,k=4)+method+s(elev, k=4)+s(plot, bs="re"), data=sm[sm$group=="wild bees",], family=poisson, method="REML"))

# syrphids
elevhv<-summary(gam(spec.no~s(elev, k=4)+s(elev,by=method,k=4)+method+s(plot, bs="re"), data=sm[sm$group=="hoverflies",], family=poisson, method="REML"))

temphv<-summary(gam(spec.no~s(tempright, k=4)+s(tempright,by=method,k=4)+method+s(elev, k=4)+s(plot, bs="re"), data=sm[sm$group=="hoverflies",], family=poisson, method="REML"))
```
